Journey: A Study in Unhelpful Titles (TODO: should we get a new title?)

### 2017-04-27 paper.tex
complete time results for mmaps

\subsubsection{Time (TODO: HP)}

Figure \ref{fig:mmap_latency} shows for various microbenchmarks the time taken per mmap operation 
as it varies with the total amount memory allocate so far in the system. As described above, we use 
rdtsc to find the number of ticks taken for the operation and then calibrate rdtsc to conver
these ticks into microseconds. It can be seen that for continuous mmaps the time is not only the lowest
but also invariant with the total number of mmaps done so far. This is because of two reasons. First, 
when a continuous mmap (with the same permissions) happens the kernel memory allocator does not need 
create a new vm area struct, instead it modifies the adjacent vm area structs to account for the newly 
accommodate memory. This ensures that there is no need to perform the lookup on the red black trees that 
store the vm area sturct so a constant line is seen.

>>>>

Figure \ref{fig:mmap_latency} shows for various microbenchmarks the time taken per mmap operation 
as it varies with the total amount memory allocate so far in the system. As described above, we use 
rdtsc to find the number of ticks taken for the operation and then calibrate rdtsc to conver
these ticks into microseconds. It can be seen that for continuous mmaps the time is not only the lowest
but also invariant with the total number of mmaps done so far. This is because of two reasons. First, 
when a continuous mmap (with the same permissions) happens the kernel memory allocator does not need 
create a new vm area struct, instead it modifies the adjacent vm area structs to account for the newly 
accommodate memory. This ensures that there is no need to perform the lookup on the red black trees that 
store the vm area sturct so a horizontal line is seen. The second important thing to note here is that 
for continuous mmap operations, the time taken is always the lowest compared to other benchmarks. This 
is also a result of expanding the existing vm area sturcts. Since the number of vm area structs is very 
low (actually just 1 for the process in question) the required node in the red black tree is always found in
the cache. For rest of the microbench marks, the kernel has to first search through the red-black tree of vm area
structs. If it finds a mapping adjacent to the new memory allocation to be done, it will extend the existing 
mapping otherwise it will create a new mapping and insert it into the red-black tree. The traversal of the 
tree is an o(long n) operation where n is the number of nodes in the tree which in this case are the vm area
mappings. This log n shape for the lookup is visible in the plots for all the benchmarks in figure \ref{fig:mmap_
latency}. The frag mmap benchmark takes the most time because for fragmented mmap, the memory is first 
allocated contiguously and then all allocated regions are expanded to a higher size. This operation needs creation
of a new memory area struct. Therefore, the kernel first looks through the red-black tree but it never finds 
a node which can be expanded to include the current mapping. So it always has to create a new mapping after 
traversing the tree and then inserting it in. It also has to delete the old mapping from the red-black tree
and then rebalnce the tree if necessary. All of these operations result in the worst case latency for mmap. 
The case of strided mmaps is very similar to frag because it will also force the kerenl to always create a new
mapping and insert it into the tree after traversing the tree first. It however does not need to delete an 
existing mapping so it is faster than the frag benchmark. Finally, rand will also search through the red-black 
tree first but it will sometimes find an existing mapping adjacent to the desired mapping which can be expanded
obviating the need to create a new vm area mapping and inserting it into red-black tree. Hence, most of the times
it is faster than stride.
